---
title: Getting Started with R
author: Emilio Luz-Ricca
output: 
  html_document:
    toc: true
    number_sections: true
---

---------

```{r setup, include = FALSE}
library(tidyverse)
library(knitr)

knitr::opts_chunk$set(fig.align = "center", warning = FALSE, message = FALSE)
```


# Producing Basic Plots with Base R

## Basic Operations

In the first part of this assignment, I use base R to produce a few simple plots.

First, I create two numeric vectors (`x` and `y`) and plot them against each other.

```{r create-vectors}
x <- 1:10
y <- 10:1

plot(x, y)
```

This plot is made much more interesting by adding additional features (i.e., title, x- and y-labels, etc.).

```{r plot-vectors}
plot(x, y, type = "b", main = "The Path of a Running Person", xlab = "Longitude", ylab = "Latitude",
     col = "maroon", pch = 3)
```

## Using a Data Frame

Now, I store the vectors in a `data.frame` object and add additional features to the plot with the `symbols` function. The idea is to use a toy dataset to demonstrate a person's path between homes in a community.

```{r path-plot}
loc_data <- data.frame(id = 1:10,
                       east = sample(1:100, 10, replace = TRUE), 
                       north = sample(1:100, 10, replace = TRUE))

symbols(loc_data$east, #adding in the squares (the homes)
        loc_data$north, 
        squares = rep(0.75, 10), 
        inches = FALSE,
        xlab = "East",
        ylab = "North")

symbols(sample(1:100, 10, replace = TRUE), #adding circles (trees..?) 
        sample(1:100, 10, replace = TRUE), 
        circles = rep(1, 10),
        inches = FALSE,
        fg = "darkgreen",
        add = TRUE)

symbols(sample(1:100, 10, replace = TRUE), #adding circles (trees..?) 
        sample(1:100, 10, replace = TRUE),
        circles = rep(1.5, 10),
        inches = FALSE,
        fg = "lightgreen",
        bg = "beige",
        add = TRUE)

to_visit <- sample(1:10, 3) #picking out the squares to visit

xspline(loc_data[to_visit, 2], #adding a path between the selected homes
        loc_data[to_visit, 3], 
        lty = 3,
        lwd = 0.75,
        shape = -1)

text(loc_data[to_visit, 2], #labeling the visited homes
     loc_data[to_visit, 3] + 4,
     labels = loc_data[to_visit, 1])

title(main = "A Person's Path Between Homes") #adding a title to the plot
```

## Challenge Question

The code for the challenge question is very similar to the previous plot, just with more data points and a more complicated path (again using `xspline`). I don't include the code for this plot as it would be repetitive. This plot is a little messy, but still somewhat readable.

```{r challenge-plot-base, echo = FALSE, out.width = "90%"}
loc_data <- data.frame(id = 1:50,
                       east = sample(1:1000, 50, replace = TRUE), 
                       north = sample(1:1000, 50, replace = TRUE))

symbols(loc_data$east, #adding in the squares (the homes)
        loc_data$north, 
        squares = rep(7, 50), 
        inches = FALSE,
        xlab = "East",
        ylab = "North")

symbols(sample(1:1000, 40, replace = TRUE), #adding circles (trees..?) 
        sample(1:1000, 40, replace = TRUE), 
        circles = rep(10, 40),
        inches = FALSE,
        fg = "darkgreen",
        add = TRUE)

symbols(sample(1:1000, 12, replace = TRUE), #adding circles (trees..?) 
        sample(1:1000, 12, replace = TRUE), 
        circles = rep(15, 12),
        inches = FALSE,
        fg = "lightgreen",
        bg = "beige",
        add = TRUE)

to_visit <- sample(loc_data$id, 7) #picking out the squares to visit

xspline(loc_data[to_visit, 2], #adding a path between the selected homes
        loc_data[to_visit, 3], 
        lty = 3,
        lwd = 0.75,
        shape = -1)

text(loc_data[to_visit, 2], #labeling the visited homes
     loc_data[to_visit, 3] + 25,
     labels = loc_data[to_visit, 1], 
     cex = 0.9)

title(main = "A Person's Path Between Homes") #adding a title to the plot
```

# Using Tidyverse and ggplot

## Reproducing the Challenge Plot

Now, I use ggplot to reproduce the challenge plot. I unfortunately couldn't figure out how to add the same shaped `xspline` to the plot, but otherwise it looks pretty similar!

```{r challenge-plot-ggplot, out.width = "80%"}
extra_points <- tibble(east = sample(1:1000, 52, replace = TRUE), #the points for the trees
                       north = sample(1:1000, 52, replace = TRUE))

for_labels <- loc_data[to_visit, ]

ggplot(loc_data, aes(x = east, y = north)) +
  theme_bw() +
  geom_point(shape = 0) + #adding the houses
  geom_point(data = extra_points[1:40, ], mapping = aes(x = east, y = north),
             size = 3.5, pch = 21, color = "darkgreen") + #adding the small trees
  geom_point(data = extra_points[41:52, ], mapping = aes(x = east, y = north),
             size = 5, pch = 21, color = "lightgreen", fill = "beige") + #adding the big trees
  geom_text(data = for_labels, mapping = aes(x = east, y = north, label = id), nudge_y = 30, fontface = "bold") +
  geom_line(data = loc_data %>% filter(id %in% to_visit), linetype = "dashed") + #visiting the desired houses
  labs(title = "A Person's Path Between Homes") + #adding a title
  theme(plot.title = element_text(hjust = 0.5)) #centering the title
```






